\documentclass[12pt,letterpaper]{article}

% packages
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
% \usepackage[utf-8]{inputenc}

% Document
\title{Decision Tree Classification Flow}
\author{Aashutosh Bista}
\date{\today}

\begin{document}

\maketitle

\section{Step 1 System Entropy}
The first step for any CART is to find the entropy of the target columns, this value is called system entropy.

\subsection{Calculations and logic assesment.}
First Single out the target columns, find out the number of classes and value counts for each class.
\begin{align*}
    a \quad &\longrightarrow \quad 200 \\
    b \quad &\longrightarrow \quad 211 \\
    c \quad &\longrightarrow \quad 192 \\
    \vdots \quad &\longrightarrow \quad \vdots \\
    n \quad &\longrightarrow \quad x
\end{align*} 

Find the probabilities for every class, say for class $a$:
\begin{align*}
    P(a) &=\frac{200}{\text{ Total Rows Number}}
\end{align*} 

Compute the system entropy using the formula: 
\begin{align*}
    H(s) &= -\sum_{i=1}^{n} P(i) \log_2(P(i))
\end{align*}

\section{Step2: Finding the Splitting criteria}
A loop is ran where every feature column is targetted and its Information Gain(I.G) is calculated, the column with higher Information gain in order are selected as priority splitting criteria.

The approach for findng I.G differs based on the type of the data the targetted feature column holds. The approach differs by evaluating the type of data on hand, if the column consists of values that can be exacted/repeated multiple times(similar to discrete), or no consistent repition(continous in nature).

\subsection{Discrete nature of Data}
These columns of data can be one hot encoded. Similar to the target column these feature columns are also supposed to have unique elements and counts associated to them. Then we perform \textbf{Binary Subset Split}. In this split as the name suggests, only two category is conisdered. Say one of the Category is $A$ then the other category becomes $~A$ or $U-A$. Then we calculate the \textbf{Weighted Entropy}. This is repeated for every Category presented in the feature column.

\subsubsection{I.G for a Discrete nature column.}
Let's assume a feature column with three distinct features in it (0,1,2) after encoding from strings and target column with target vairables as (A,B,C). First we compute, the metrics such as counts of mapping such as counts of a feature varaible and target variable.

\begin{align*}
    0_A \quad &\longrightarrow \quad 344 \\
    1_A \quad &\longrightarrow \quad 221 \\
    2_A \quad &\longrightarrow \quad 343 \\
    0_B \quad &\longrightarrow \quad 222 \\
    \vdots \quad &\longrightarrow \quad \vdots \\
    2_C \quad &\longrightarrow \quad 212
\end{align*}
Now, lets compute the binary split of $0$ and $0^{\complement}$.

we have $0$ and $0^{\complement}$, our goal is to find the Weighted entropy of a binary split, so the formual becomes.
\begin{align*}
    H_w(X) &= \sum_{i=1}^{n} P(0_i) H(0_i) + P(0^{\complement}_i)H(0^{\complement}_i) 
\end{align*}

The probabilities are used as weights and the notations are tried to be summarized below:
\begin{align*}
    n = \text{Number of Unique classes in target column} \\
    P(0_i) = \text{weight 1 (Probability of 0 in feat col)}\\
    P(0^{\complement}_i) = \text{ weight 2 (Probability of $0^{\complement}$ in feat col)}\\
    H(0_i) = \text{Entropy of 0th component of binary split}\\
    H(0^{\complement}_i) = \text{Entropy of 1th component of binary splilt} 
\end{align*}

Entropy of the binary splits:
\begin{align*}
    H(0_i) &= P(0_A) \log2(P(0_A))+P(0_B) \log2(P(0_B))+P(0_C) \log2(P(0_C)) \\
    H(0^{\complement}_i) &= P(0^{\complement}_A) \log2(P(0^{\complement}_A))+P(0^{\complement}_B) \log2(P(0^{\complement}_B))+P(0^{\complement}_C) \log2(P(0^{\complement}_C))
\end{align*}
We repeate this for the set ($1$ ,$1^{\complement}$) and ($2$, $2^{\complement}$), and find the I.G by subtracting the weighted entropy for each set.


\subsection{Continuos nature of Data}
To find the splitting criteria in a column of continous nature modern CART models use a combined approach of \textbf{Greedy Approach and Histogram Binning}. Let's look at them individually.

\subsubsection{Greedy Approach}
In Greedy Approach, we leave out no possible candidate for splitting, this is definitely a compute heavy approach. Lets imagine a continous feature columns and it's corresponding target column with three specific targets $A$,$B$ and $C$.

\begin{table}[H]
    \centering
    \label{Tabular view of distribution}
    \begin{tabular}{|c|c|}
        \hline
        Feat Col & Targ Col \\
        \hline
        21 & B \\
        14 & C \\
        42 & C \\
        19 & A \\
        36 & A \\
        21 & B \\
        14 & A \\
        59 & B \\
        68 & C \\
        75 & C \\
        \hline
    \end{tabular}
    \caption{Feat and Targ cols}
\end{table}

This apporach is supposed to work as, find the mean between every sorted unique elements in the feature columns and find the Information Gain for that mean and find the one with the Highest Information Gain and use it as the splitting criteria. 

\begin{table}[H]
    \centering
    \label{Tabular view}
    \begin{tabular}{|c|c|c|}
        \hline
        Sorted Feats & Targ Col & Upper and Lower end\\
        \hline
        14 & A & --\\
        14 & C & 14-19\\
        19 & A & 19-21\\
        21 & B & --\\
        21 & B & 21-36\\
        36 & A & 36-42\\
        42 & C & 42-59\\
        59 & B & 59-68\\
        68 & C & 68-75\\
        75 & C & Upper right end \\
        \hline
    \end{tabular}
    \caption{Feat and Targ cols}
\end{table}
We obviously ignore the repeated values as they serve as the same bounding parameter while calculating the mean, but the associated value in the target column should be recorded for entropy calculation.

\begin{itemize}
    \item \textbf{Lopping} \newline
    We run the loop $k-1$, where $k$ is the number of rows in the given data.
    \item \textbf{Entropy Calculation} \newline
    For every mean we look for the 
    \begin{align*}
         H_L &= \sum P(A_L) \log2(P(A_L)) + P(B_L) \log2(P(B_L)) + P(C_L) \log2(P(C_L)) \\
         \text{and} \\
         H_R &= \sum P(A_R) \log2(P(A_R)) + P(B_R) \log2(P(B_R)) + P(C_R) \log2(P(C_R)) \\
    \end{align*}
    where,
    \begin{align*}
    P(A_L) = \text{Porbability of A in left side of mean} \\
    P(B_L) = \text{Porbability of B in left side of mean}\\
    P(C_L) = \text{Porbability of C in left side of mean}\\
    P(A_R) = \text{Porbability of A in right side of mean} \\
    P(B_R) = \text{Porbability of B in right side of mean}\\
    P(C_R) = \text{Porbability of C in right side of mean}\\
    \end{align*}
    \item Weighted Entropy \newline
    \begin{align*}
    H_w(X) &= \sum P(L) H(L) + P(R)H(R) 
\end{align*}
    where,
    \begin{align*}
    P(L) = \text{Porbability of a entity being on left of mean} \\
    P(R) = \text{Porbability of a entity being on right of mean}\\
    \end{align*}
\end{itemize}

\section{Testing column mapping dynamically}
    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|c|}
            \hline
            0 & 1 & 2 & 3 & 4 & 5 & 6 & 7\\
            \hline
            A & B & C & D & E & F & G & H\\
            \hline
        \end{tabular}
        \caption{primary state depth 0}
    \end{table}
    \begin{align*}
        \text{Gone}\\
        3 \quad &\longrightarrow \quad D\\
        depth = 0\\
        \text{Actual}: None
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}


    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 & 1 & 2 & 3 & 4 & 5 & 6 \\
            \hline
            A & B & C & E & F & G & H\\
            \hline
        \end{tabular}
        \caption{primary state depth 1}
    \end{table}
    \text{1:B}
        \begin{align*}
        \text{Gone}\\
        1 \quad &\longrightarrow \quad B\\
        \text{Actual(1:B)}\\
        depth = 1 \\
        col_{prev} = 3 \quad col_{curr} = 1 \\
        col_{prev} > col_{curr}\\
        \text{}
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}

    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 & 1 & 2 & 3 & 4 & 5\\
            \hline
            A & C & E & F & G & H\\
            \hline
        \end{tabular}
        \caption{primary state depth 2}
    \end{table}
    \text{5:H}
    \begin{align*}
        \text{Gone}\\
        5 \quad &\longrightarrow \quad H\\
        \text{Actual(7:H)}\\
        depth = 2 \\
        col_{prev} = 1 \quad col_{curr} = 5 \\
        col_{prev} < col_{curr}\\
        \text{}
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}

    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 & 1 & 2 & 3 & 4\\
            \hline
            A & C & E & F & G\\
            \hline
        \end{tabular}
        \caption{primary state depth 3}
    \end{table}
    \text{2:E}
    \begin{align*}
        \text{Gone}\\
        2 \quad &\longrightarrow \quad E\\
        \text{Actual(4:E)}\\
        depth = 3 \\
        col_{prev} = 5 \quad col_{curr} = 4 \\
        col_{prev} > col_{curr}\\
        \text{}
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}

    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 & 1 & 2 & 3\\
            \hline
            A & C & F & G\\
            \hline
        \end{tabular}
        \caption{primary state depth 3}
    \end{table}
    \text{0:A}
    \begin{align*}
        \text{Gone}\\
        0 \quad &\longrightarrow \quad A\\
        \text{Actual(0:A)}\\
        depth = 4 \\
        col_{prev} = 4 \quad col_{curr} = 0 \\
        col_{prev} > col_{curr}\\
        \text{}
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}


    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 & 1 & 2\\
            \hline
            C & F & G\\
            \hline
        \end{tabular}
        \caption{primary state depth 4}
    \end{table}
    \text{1:F}
    \begin{align*}
        \text{Gone}\\
        1 \quad &\longrightarrow \quad F\\
        \text{Actual(5:F)}\\
        depth = 5 \\
        col_{prev} = 0 \quad col_{curr} = 5 \\
        col_{prev} < col_{curr}\\
        \text{}
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}

    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 & 1\\
            \hline
            C & G\\
            \hline
        \end{tabular}
        \caption{primary state depth 5}
    \end{table}
    \text{0:C}
    \begin{align*}
        \text{Gone}\\
        0 \quad &\longrightarrow \quad C\\
        \text{Actual(2:C)}\\
        depth = 6 \\
        col_{prev} = 5 \quad col_{curr} = 0 \\
        col_{prev} < col_{curr}\\
        \text{}
    \end{align*}
\noindent\rule{\linewidth}{0.4pt}


    \begin{table}[H]
        \centering
        \label{Tabular view of distribution}
        \begin{tabular}{|c|c|c|c|c|c|c|}
            \hline
            0 \\
            \hline
            G\\
            \hline
        \end{tabular}
        \caption{primary state depth 6}
    \end{table}
    % \text{1:F}

    iter 0 
    [3]
    iter 1 []


\subsubsection{Histogram Binning}
This is the more of compute friendly approach where, we divide the data in bins and try singling out one of the bin, which is able to point out the maximum information for that column, the bins is either averaged or one of the end is taken as the criteria. This apporach is also called \textbf{Histogram Binning}.


\end{document}